---
title: "Veiledning i SSBpris"
author: "Anne Vedø"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SSBpris-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
```

### Introduction

The package SSBpris is used for calculating price indexes and their variance. It currently contains two functions:

* CalcInd - For calculating estimates of price indexes
* CalcIndS2 - For calculating estimates of the variance of price indexes

The package handles three types of elementary indexes: Jevons, Dutot and Carli

### Data requirements

The data set must contain the following variables:

* Price in base period 
* Price in statistical time period
* Elementary group 
* A weight for aggregating elementary groups

Optional variable:

* A variable for grouping elementary groups, used for calculating subindexes.


### Examples

Using CalcInd without consumVar gives all the elementary indexes as output.   

```{r} 
library(SSBpris)
data("priceData")

CalcInd(data=priceData, baseVar="b1", pVar="p1", type="jevons", groupVar="varenr", 
        consumVar = NULL, wVar = "weight") 
```
   
Using CalcInd with consumVar gives an index for each level of consumVar as output.   

```{r} 
CalcInd(data=priceData, baseVar="b1", pVar="p1", type="jevons", groupVar="varenr", 
        consumVar = "coicop", wVar = "weight") 
```                  
                   